---
title: "Jointly estimate effective reproduction number and gravity model parameters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simulating-data}
  %\VignetteEngine{rmarkdown::render}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(ollapse = TRUE, comment = "#>")
```

```{r setup}
library(EpiEstim)
library(geosphere)
library(ggplot2)
library(MASS)
library(spatialbranchr)
library(tidyr)
```

We will data from the West African Ebola epidemic to jointly estimate
the effective reproduction number and two parameters of the gravity
model capturing human population movement - (a) pstay, which characterises the probability that an
infectious individual will not move out of a spatial unit (here
country) during their infectious period, and (b) gamma, which
modulates the effect of the distance between two spatial units on the
flow of populations between them.


## Data

We will use 

```{r}
incid <- readr::read_csv("vignettes/promed_loglinear_wide_from_march_ghana_fixed.csv")
## Get rid of dates column as we don't need it for fitting
incid_raw <- incid[, -1]
nloc <- ncol(incid_raw)
centroids <- readr::read_csv("vignettes/all_african_centroids.csv")
distances <- readRDS("vignettes/allafrica_distances.rds")
si <- EpiEstim::discr_si(k = 0:35, mu = 6, sigma = 4)
priors <- spatial_priors()

cntrs <- c("SLE", "GIN", "LBR", "GHA", "NGA")
x <- incid_raw[1:35, cntrs]
pop <- centroids$Pop[centroids$ISO3 %in% cntrs]
coords <- centroids[centroids$ISO3 %in% cntrs, c("Centroid_Lon", "Centroid_Lat")]
dist <- distm(coords)
out <- spatial_estimate(
  x, si, 7, NULL, pop, dist, 1, 1, 1, "poisson", priors
)

```
