---
title: "Simulating outbreaks under a spatially explicit branching process"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simulating-data}
  %\VignetteEngine{rmarkdown::render}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(EpiEstim)
library(ggplot2)
library(MASS)
library(spatialbranchr)
library(tidyr)
```

The model implemented in `spatialbranchr` 
relies on a well-established statistical framework that
assumes the daily incidence, \(I_t\) , can be approximated with a
Poisson process following the renewal equation \cite{fraser2007estimating}:

\begin{equation}
I_{t} \sim 
  Pois\left( 
  R^{t} \sum_{s = 1}^{t}{I_{i}^{t - s} \omega_{s}}
  \right).
  \label{eq:likelihood1}
\end{equation}

Here \(R^t\) is the reproduction number at time \(t\) (the average
number of secondary cases per infected individual) and \(\omega\) is the
distribution of the serial interval (the time between onset of symptoms
in a case and their infector).

We extend this model to incorporate the spatial spread of the outbreak
between \(n\) different locations. We assume that the number of incident cases at a
location $j$ at time $t$ is given by the equation

\begin{equation}
I_{j}^{t} \sim Pois\left( \Lambda_{j}^{t} \right), 
\label{eq:likelihood2}
\end{equation}

where $\Lambda_{j}^{t}$ is the total infectivity at a location $j$ at
time $t$. $\Lambda_{j}^{t}$ is
the sum of infectivity at all locations weighted by the relative flow of
cases into $j$ from each location $i$. That is, 

\begin{equation}
\Lambda_{j}^{t} = \sum_{i = 1}^{n} 
  {\left( p_{i \rightarrow j}
  R_{i}^{t} \sum_{s = 1}^{t}{I_{i}^{t - s} \omega_{s}}
  \right)}.
\label{eq:biglambda}
\end{equation}

$R_{i}^{t}$ is the reproduction number at location $i$ at time $t$
and $p_{i \rightarrow j}$ is the probability of a case moving from
location $i$ to location $j$ while they are infectious. \(\omega\) is
the typical infectiousness profile of a case over time after
infection. 

We illustrate the model by simulating incidence for three locations
under various assumptions of population movement between the three places.

## No movement

If we don't allow any movement between the three locations, we are in
effect simulating three independent outbreaks.

```{r}
locations <- c("A", "B", "C")
## Movement matrix, all off-diagonal elemnets are 0.
pmat <- matrix(0, nrow = 3, ncol = 3)
diag(pmat) <- 1
## Initial Rt in the three locations
r_init <- matrix(rep(2, 3), nrow = 1, ncol = 3)
## Initial incidence, start with a large seed so that the epidemic can
## take off.
incid_init <- matrix(rep(20, 3), nrow = 1, ncol = 3)

si <- EpiEstim::discr_si(k = 0:30, mu = 6, sigma = 4)

no_movement <- spatial_project(
  incid_init, r_init, si, pmat, n_sim = 1000, n_days = 30,
  model = "poisson"
)

no_movement_df <- as_dataframe(no_movement)
tall <- gather(no_movement_df, location, incid, -sim, -day)

```

```{r}

ggplot(tall) +
  geom_line(aes(day, incid, group = sim), alpha = 0.3) +
  facet_wrap(~location, nrow = 3)

```

## Allow movement

We will now modify the probability of movement to allow for some
movement between locations.

```{r}
puni <- matrix(
  c(0.8, 0.1, 0.1, 0.2, 0.3, 0.5, 0.3, 0.3, 0.4), nrow = 3, ncol = 3
)
uni_movement <- spatial_project(
  incid_init, r_init, si, puni, n_sim = 1000, n_days = 30,
  model = "poisson"
)

uni_movement <- as_dataframe(uni_movement)
tall_uni <- gather(uni_movement, location, incid, -sim, -day)

```

```{r}
ggplot(tall) +
  geom_line(aes(day, incid, group = sim), alpha = 0.3) +
  facet_wrap(~location, nrow = 3)

```
